---
title: "Homework 9"
author: "Arjun Yenamandra - SDS 315 UT Austin"


output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=7, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```
eid: ay7654
github: [https://github.com/macksed/HW9](https://github.com/macksed/HW9)
```{r, echo=FALSE, results='hide'}
library(tidyverse)
library(ggplot2)
library(mosaic)
library(knitr)
```
# **Question 1**
### **A**
```{r, echo=FALSE, results='hide'}
turnout <- read.csv("turnout.csv")
```

```{r, echo=FALSE, results='hide'}
votedg <- sum(turnout$voted1998[turnout$GOTV_call == 1])
totalg <- length(turnout$voted1998[turnout$GOTV_call == 1])
votednog <- sum(turnout$voted1998[turnout$GOTV_call == 0])
totalnog <- length(turnout$voted1998[turnout$GOTV_call == 0])

# Use prop.test to calculate the test of proportions and confidence interval
test_result <- prop.test(x = c(votedg, votednog), 
                         n = c(totalg, totalnog),
                         correct = TRUE) # using continuity correction

# Printing the test results, which include confidence interval
(test_result)
```
```{r, echo=FALSE}
(test_result)
```


### **B**


```{r, echo=FALSE, results='hide'}
summary_stats <- turnout %>%
  group_by(GOTV_call) %>%
  summarise(
    MAge = mean(AGE, na.rm = TRUE),
    Prop_MAJORPTY = mean(MAJORPTY, na.rm = TRUE),
    Prop_1996 = mean(voted1996, na.rm = TRUE)
  )
s = as_tibble(summary_stats)
s
votedvsGOTV <- chisq.test(turnout$voted1996, turnout$GOTV_call)
print(votedvsGOTV)

# Association between AGE and GOTV_call
AGOTV <- t.test(turnout$AGE ~ turnout$GOTV_call)
print(AGOTV)

# Association between MAJORPTY and GOTV_call
MAJORPTYG <- chisq.test(turnout$MAJORPTY, turnout$GOTV_call)
print(MAJORPTYG)

v96v98 <- chisq.test(turnout$voted1996, turnout$voted1998)
print(v96v98)

# Association between AGE and voted1998
A1998 <- t.test(turnout$AGE ~ turnout$voted1998)
print(A1998)

# Association between MAJORPTY and voted1998
MAJORPTY98 <- chisq.test(turnout$MAJORPTY, turnout$voted1998)
print(MAJORPTY98)

print(turnout %>%
  group_by(GOTV_call, voted1996, voted1998, MAJORPTY) %>%
  summarise(count = n()))
```
If we calculate the chisquared value for each categorical confounder against both 1998 voting (outcome)and if one was called(treatment), we see that every single value shows a clear difference, all with p values that are extremely small. All of the p values were < 2e-5, except one, at 0.505. For age, we conduct a confidence interval and we find that those who did receive a call were older, and those who did vote were older. Both of these with similar differentials under the 95% confidence interval.

```{r, echo=FALSE, results='hide'}
ggplot(turnout, aes(x = as.factor(voted1996), fill = as.factor(GOTV_call))) +
  geom_bar(position = "dodge") +
  labs(x = "Voted in 1996", fill = "GOTV Call")
```

```{r, echo=FALSE, results='hide'}
ggplot(turnout, aes(x = as.factor(MAJORPTY), fill = as.factor(GOTV_call))) +
  geom_bar(position = "dodge") +
  labs(x = "Major Party", fill = "GOTV Call")
```
If we look at the proportional differences in party and if people voted or not, we can see that there is a clear discrepancy among our confounding variables. After some visualization, we can confirm that all three are confounding variables.
```{r, echo=FALSE, results='hide'}
library(MatchIt)
```

### **C**
```{r, echo=FALSE, results='hide'}
matcher <- matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, data = turnout, method = "nearest", ratio = 5)
```

```{r, echo=FALSE, results='hide'}
matched_data <- match.data(matcher)
votedg_matched <- sum(matched_data$voted1998[matched_data$GOTV_call == 1])
totalg_matched <- length(matched_data$voted1998[matched_data$GOTV_call == 1])
votednog_matched <- sum(matched_data$voted1998[matched_data$GOTV_call == 0])
totalnog_matched <- length(matched_data$voted1998[matched_data$GOTV_call == 0])
```

```{r, echo=FALSE, results='hide'}
test_result_matched <- prop.test(x = c(votedg_matched, votednog_matched), 
                                 n = c(totalg_matched, totalnog_matched),
                                 correct = TRUE)  
print(test_result_matched)



votedvsGOTV <- chisq.test(matched_data$voted1996, matched_data$GOTV_call)
print(votedvsGOTV)


AGOTV <- t.test(matched_data$AGE ~ matched_data$GOTV_call)
print(AGOTV)

MAJORPTYG <- chisq.test(matched_data$MAJORPTY, matched_data$GOTV_call)
print(MAJORPTYG)

v96v98 <- chisq.test(matched_data$voted1996, matched_data$voted1998)
print(v96v98)

A1998 <- t.test(matched_data$AGE ~ matched_data$voted1998)
print(A1998)

MAJORPTY98 <- chisq.test(matched_data$MAJORPTY, matched_data$voted1998)
print(MAJORPTY98)

print(matched_data %>%
  group_by(GOTV_call, voted1996, voted1998, MAJORPTY) %>%
  summarise(count = n()))
```
I believe there is significant evidence that the call increased the odds of voting in 1998. The matching allowed for age to be normalized, and proved that when the proportions of 96 voters and party was non confounding, the call had a proven effect.

# **Question 2**

### **A**

```{r, echo=FALSE, results='hide'}
solder <- read.csv("solder.csv")
```

```{r, echo=FALSE, results='hide'}
 ggplot(solder, aes(x = Opening, y = skips)) +
  geom_boxplot() +
  labs(title = "Relationship between Opening Size and Number of Skips",
       x = "Opening Size",
       y = "Number of Skips")

```





This plot shows that as the opening size gets smaller, the number of skips actually increases.
```{r, echo=FALSE, results='hide'}
ggplot(solder, aes(x = Solder, y = skips)) +
  geom_boxplot() +
  labs(title = "Relationship between Solder Thickness and Number of Skips",
       x = "Solder Thickness",
       y = "Number of Skips")

```





This plot shows that as the solder gets thinner, the number of skips increases.The distribution is shifted higher, and the quartiles/median are all generally higher.



### **B**
```{r, echo=FALSE, results='hide'}
skipsmodel = lm(skips ~ Opening*Solder, data = solder)
conf_intervals <- confint(skipsmodel, level = 0.95)
estimates  = summary(skipsmodel)$coefficients[,1]
summary(skipsmodel)
coefs = c('Intercept','OpeningM','OpeningS','SolderThin','OpeningM:SolderThin','OpeningS:SolderThin')

tables = as_tibble(cbind(coefs,conf_intervals,estimates))

```

```{r, echo=FALSE}
print(kable(tables))
```
This shows our 95% confidence interval, and our estimates.



### **C**
The baseline skips for a large opening, and thick solder is 0.3933.
The main effect for the FarAway variable is 31 ms. This is the effect of FarAway in isolation.
The main effect for the Opening variable in OpeningM is 2.4067 skips. The main effect for the Opening variable in OpeningS is 5.1267 skips. The main effect for the Solder variable in SolderThin is 2.28 skips. 
The interaction effect for OpeningM and SolderThin is -.74 skips. In other words, both medium opening and thin solders are -.74 skips less than what you would expect from summing the individual “isolated” effects of the two variables. In the same light, the interaction for OpeningS and thin solders is 9.6533. Thus they cause ~9.7 more skips than isolated.




### **D**
OpeningL and SolderThick would minimize the amount of skips. with this combination, our only value is the value of the intercept, as all the other interactions get multiplied by 0 as the condition is not met. For the other coefficients, we have to add multiple positive values for the expected skips. This surely is larger than the intercept, and thus OpeningL and SolderThick is the best combination.

```{r, echo=FALSE, results='hide'}

```

```{r, echo=FALSE, results='hide'}

```

```{r, echo=FALSE, results='hide'}

```

```{r, echo=FALSE, results='hide'}

```

```{r, echo=FALSE, results='hide'}

```

```{r, echo=FALSE, results='hide'}

```

```{r, echo=FALSE, results='hide'}

```











